Manuscript title here

### 2024-09-16 manuscript.Rmd
la till tanke idé för ofi och chock koppling. nu äntligen fattar jag typ vad som ska vara med i intro

Trauma is a major global public health concern. It causes over four milion death and affects hundreds of millions each year, among those younger populations are the most affected [@noauthor_injuries_nodate]. Shock is a major cause of mortality in trauma patients. The definition for shock is circulatory failure leading to insuficient perfusion and oxygenation of tissue and organ[@leech_shock_2023]. There are four types of causes, which are hypovolemic, obstructive, cardiogenic and distributive. These causes can occur alone or in combination. The most usual one during trauma are hypovolemic in form of a major haemorrhage [@teixeira_preventable_2007].

@@

The initial management of trauma patients once they reach hospital is very time sensitive and error prone [@mackersie_pitfalls_2010]. To improve trauma care quality, healthcare systems identify Opportunities for Improvement (OFI) through structured reviews of patient cases. Multidisciplinary morbidity and mortality reviews play a central role in this process, where clinicians from different specialities together analyse trauma cases to evaluate both the structure and clinical processes involved in trauma care following the Donabedian’s quality of care framework[@albaaj_patient_2023]. For shock patients, timely interventions such as fluid resuscitation/blood transfusion, intervention, OR and radiology are critical, and delays or missteps can result in adverse outcomes (källa). By systematically addressing OFIs, healthcare providers can improve the effectiveness of trauma care, directly impacting mortality and morbidity.

>>>>

Trauma is a major global public health concern. It causes over four milion death and affects hundreds of millions each year, among those younger populations are the most affected. Not only does it do bodily harm, but also places a burden on national economies, in form of billions of us dollars for health care, lost productivity and law enforcement[@noauthor_injuries_nodate].

Shock is a major cause of mortality in trauma patients[@siegel_effect_1995]. The definition for shock is circulatory failure leading to insuficient perfusion and oxygenation of tissue and organ[@leech_shock_2023]. There are four types of causes, which are hypovolemic, obstructive, cardiogenic and distributive. These causes can occur alone or in combination. The most usual one during trauma are hypovolemic in form of a major haemorrhage [@teixeira_preventable_2007].

@@

The initial management of trauma patients once they reach hospital is very time sensitive and error prone [@mackersie_pitfalls_2010].

This is especially worth noting when patients in shock also have a higher mortality rate than those not[@vang_shock_2022].

Koppla ihop det mpå nått sätt med att, trauma i sig är tid sensitiv och kaotisk, kombinerat med att chock patienter har högre mortalitet (kanske källa som även säger chock pat är mer kaotisk och tid sensitiv), parallellt har pat med medel+hög iss hög OFI, dessa medel och hög har oftast chock(källa), därför kombinerat kaos natur + kaos från chock + medel skadade oftast med chock har hög ofi -> bör undersöka pat med olika nivåer av chock har ofi.

To improve trauma care quality, healthcare systems identify Opportunities for Improvement (OFI) through structured reviews of patient cases. Multidisciplinary morbidity and mortality reviews play a central role in this process, where clinicians from different specialities together analyse trauma cases to evaluate both the structure and clinical processes involved in trauma care following the Donabedian’s quality of care framework[@albaaj_patient_2023]. For shock patients, timely interventions such as fluid resuscitation/blood transfusion, intervention, OR and radiology are critical, and delays or missteps can result in adverse outcomes (källa). By systematically addressing OFIs, healthcare providers can improve the effectiveness of trauma care, directly impacting mortality and morbidity.
